Variants of G\"{o}del's Ontological Proof \\ in a Natural Deduction Calculus

### 2016-08-15 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
(D.6) Section 4

\begin{definition} \emph{Terms and formulas} are given by the following grammar:
\begin{align*}
 s,t \quad ::= \quad & 
  p_\tau \ \mid \ 
  X_\tau \ \mid \
  (\lambda X_\theta.s_\tau)_{\theta\imp\tau} \ \mid \ 
  (s_{\theta\imp\tau}\, t_\theta)_\tau \ \mid \\
& \bot_o \ \mid \
  \imp_{o\imp o\imp o} \ \mid \ 
  \wedge_{o\imp o\imp o} \ \mid \
  \vee_{o\imp o\imp o} \ \mid \\
& \all_{(\tau\imp o)\imp o} \ \mid \ 
  \ex_{(\tau\imp o)\imp o} \ \mid \
  \nec_{o\imp o} \ \mid \
  \pos_{o\imp o}
\end{align*}
where $p_\tau$ and $X_\tau$ range over, respectively, constants and variables of type $\tau$. Parenthesis conventions, infix notation for propositional connectives and binding notation for quantifiers are assumed. Furthermore, subscript types are omitted when they are clear from the context. Negation ($\neg_{o\imp o}$) and equivalence ($\biimp_{o\imp o\imp o}$) are defined by $\neg A\equiv A\imp \bot$ and $ (A\biimp B)\equiv (A\imp B)\wedge (B\imp A)$.
\end{definition}

@@

\begin{small}
\begin{prooftree}
		\AXC{$\subproof_2$} \dashedLine
		\UIC{$P(\lambda x. x \neq x)\scott$}    
    		\AXC{$\subproof_1$} \dashedLine
			  \UIC{$P(\lambda x. x = x)\scott$} 
					\AXC{Half of Axiom 1} \dashedLine
					\UIC{$P(\lambda x. x = x) \imp \neg P(\lambda x. x \neq x)$} \RightLabel{$\imp_E$}
				\BIC{$\neg P(\lambda x. x \neq x)$} \RightLabel{$\neg_E$}
			\BIC{$\bot$}
\end{prooftree}
\end{small}

\begin{small}
\begin{prooftree}
\AXC{$ $} \RightLabel{4}
\UIC{$\nec \neg \ex x. \rho(x)$} \RightLabel{$\nec_E$}
\UIC{$\omega: \neg \ex x. \rho(x)$}
        \AXC{$ $} \RightLabel{2}
        \UIC{$\rho(\alpha)$} \RightLabel{$\ex_I$}
		\UIC{$\omega: \ex x. \rho(x)$} \RightLabel{$\neg_E$}
    \BIC{$\omega: \bot$} \RightLabel{$\neg_I^2$}
	\UIC{$\omega: \neg \rho(\alpha)$} \RightLabel{$\all_I$}
	\UIC{$\omega: \all x. \neg \rho(x)$} \RightLabel{$\nec_I$}
	\UIC{$\nec \all x. \neg \rho(x) \scott$}
	        \AXC{$ $} \dashedLine
	        \UIC{$\bot$} \RightLabel{$\neg_I^3$}
    		\UIC{$\neg \nec \all x. \neg \rho(x)$} \RightLabel{$\neg_E$}
  		\BIC{$\bot$} \RightLabel{$\neg_I^4$}
  		\UIC{$ \neg \nec \neg \ex x.\rho(x)$} \doubleLine %\RightLabel{definition of $\pos$}
  		\UIC{$ \pos \ex x.\rho(x)$}  \RightLabel{$\imp_I^5$}
  		\UIC{$ P(\rho) \imp \pos \ex x.\rho(x)$} \RightLabel{$\all_I$}
      \UIC{$ \all \varphi. [P(\varphi) \imp \pos \ex x.\varphi(x)] $}
\end{prooftree}
\end{small}
\end{proof}

@@

					\AXC{$ $} \RightLabel{5}
					\UIC{$G(\gamma)$} \dottedLine
					\UIC{$\all \varphi. P(\varphi) \imp \varphi(\gamma)$} \RightLabel{$\all_E$}
					\UIC{$P(\lambda x. \neg \rho(x)) \imp \neg \rho(\gamma)$} \RightLabel{$\imp_E$}
				\BIC{$\neg \rho(\gamma)$} \RightLabel{$\neg_E$}      
	   \BIC{$\bot$} \RightLabel{$\neg_I^1$}
	   \UIC{$\neg\neg P(\rho)$} \RightLabel{$\neg\neg_E$}
	   \UIC{$[\subproof_3] \qquad P(\rho)$}
	   \end{prooftree}
\end{scriptsize}

@@

 \AXC{$\subproof_3$}\dashedLine
 \UIC{$P(\rho)$}
	      \AXC{Axiom 4} \dashedLine
        \UIC{$\all \varphi.[P(\varphi) \to \Box \; P(\varphi)]$} \RightLabel{$\all_E$}
	   	  \UIC{$P(\rho) \imp \nec P(\rho)$} \RightLabel{$\imp_E$}
		 \BIC{$\nec P(\rho)\scott$}
			\AXC{$ $} \RightLabel{3}
			\UIC{$\omega: P(\rho)$}
					\AXC{$ $} \RightLabel{2}
					\UIC{$\omega: G(\delta) $} \dottedLine
					\UIC{$\omega: \all \varphi. P(\varphi) \imp \varphi(\delta)$} \RightLabel{$\all_E$} 
					\UIC{$\omega: P(\rho) \imp \rho(\delta)$} \RightLabel{$\imp_E$}
				\BIC{$\omega: \rho(\delta) $} \RightLabel{$\imp_I^2$}
				\UIC{$\omega: G(\delta) \imp \rho(\delta) $} \RightLabel{$\all_I$}
				\UIC{$\omega: \all y. G(y) \imp \rho(y) $} \RightLabel{$\imp_I^3$}
				\UIC{$\omega: P(\rho) \imp (\all y. G(y) \imp \rho(y) )$} \RightLabel{$\nec_I$}
				\UIC{$\nec (P(\rho) \imp (\all y. G(y) \imp \rho(y) ))$} \RightLabel{$\imp_E$}
						\AXC{Axiom K} \RightLabel{$\imp_E$}
					\BIC{$\nec P(\rho) \imp \nec (\all y. G(y) \imp \rho(y) )$} \RightLabel{$\imp_E$}
			\BIC{$\nec (\all y. G(y) \imp \rho(y) )$} \RightLabel{$\imp_I^4$}
			\UIC{$\rho(\gamma) \imp \nec (\all y. G(y) \imp \rho(y) )$} \RightLabel{$\all_I$}
			\UIC{$\all \varphi. \varphi(\gamma) \imp \nec (\all y. G(y) \imp \varphi(y) )$} \RightLabel{$\wedge_I$}
		\BIC{$ G(\gamma) \wedge (\all \varphi. \varphi(\gamma) \imp \nec (\all y. G(y) \imp \varphi(y) ))$} \dottedLine
		\UIC{$\ess{G}{\gamma}$} \RightLabel{$ \imp_I^5$}
		\UIC{$G(\gamma) \imp \ess{G}{\gamma}$} \RightLabel{$\all_I$}
		\UIC{$\all x. G(x) \imp \ess{G}{x}$}
\end{prooftree}
\end{scriptsize}

>>>>

\begin{definition} \emph{Terms and formulas} are given by the following grammar:
\begin{align*}
 s,t \quad ::= \quad & 
  p_\tau \ \mid \ 
  X_\tau \ \mid \
  (\lambda X_\theta.s_\tau)_{\theta\imp\tau} \ \mid \ 
  (s_{\theta\imp\tau}\, t_\theta)_\tau \ \mid \\
  & s_{\tau} = t_{\tau} \ \mid \\
& \bot_o \ \mid \
  \imp_{o\imp o\imp o} \ \mid \ 
  \wedge_{o\imp o\imp o} \ \mid \
  \vee_{o\imp o\imp o} \ \mid \\
& \all_{(\tau\imp o)\imp o} \ \mid \ 
  \ex_{(\tau\imp o)\imp o} \ \mid \
  \nec_{o\imp o} \ \mid \
  \pos_{o\imp o}
\end{align*}
where $p_\tau$ and $X_\tau$ range over, respectively, constants and variables of type $\tau$. Parenthesis conventions, infix notation for propositional connectives and binding notation for quantifiers are assumed. Furthermore, subscript types are omitted when they are clear from the context. Inequality ($\neq_{\tau\imp\tau\imp o}$), negation ($\neg_{o\imp o}$) and equivalence ($\biimp_{o\imp o\imp o}$) are defined by $s \neq t\equiv \neg (s=t)$, $\neg A\equiv A\imp \bot$ and $ (A\biimp B)\equiv (A\imp B)\wedge (B\imp A)$.
\end{definition}

@@

\begin{small}
\begin{prooftree}
		\AXC{$\subproof_2$} \dashedLine
		\UIC{$P(\lambda x. x \neq x)\scott$}    
    		\AXC{$\subproof_1$} \dashedLine
			  \UIC{$P(\lambda x. x = x)\scott$} 
					\AXC{Half of Axiom 1} \dashedLine
					\UIC{$P(\lambda x. x = x) \imp \neg P(\lambda x. x \neq x)$} \RightLabel{$\imp_E$}
				\BIC{$\neg P(\lambda x. x \neq x)$} \RightLabel{$\neg_E$}
			\BIC{$[\subproof_3] \qquad \bot$}
\end{prooftree}
\end{small}

\begin{small}
\begin{prooftree}
\AXC{$ $} \RightLabel{4}
\UIC{$\nec \neg \ex x. \rho(x)$} \RightLabel{$\nec_E$}
\UIC{$\omega: \neg \ex x. \rho(x)$}
        \AXC{$ $} \RightLabel{2}
        \UIC{$\omega: \rho(\alpha)$} \RightLabel{$\ex_I$}
		\UIC{$\omega: \ex x. \rho(x)$} \RightLabel{$\neg_E$}
    \BIC{$\omega: \bot$} \RightLabel{$\neg_I^2$}
	\UIC{$\omega: \neg \rho(\alpha)$} \RightLabel{$\all_I$}
	\UIC{$\omega: \all x. \neg \rho(x)$} \RightLabel{$\nec_I$}
	\UIC{$\nec \all x. \neg \rho(x) \scott$}
	        \AXC{$\subproof_3 $} \dashedLine
	        \UIC{$\bot$} \RightLabel{$\neg_I^3$}
    		\UIC{$\neg \nec \all x. \neg \rho(x)$} \RightLabel{$\neg_E$}
  		\BIC{$\bot$} \RightLabel{$\neg_I^4$}
  		\UIC{$ \neg \nec \neg \ex x.\rho(x)$} \doubleLine %\RightLabel{definition of $\pos$}
  		\UIC{$ \pos \ex x.\rho(x)$}  \RightLabel{$\imp_I^5$}
  		\UIC{$ P(\rho) \imp \pos \ex x.\rho(x)$} \RightLabel{$\all_I$}
      \UIC{$ \all \varphi. [P(\varphi) \imp \pos \ex x.\varphi(x)] $}
\end{prooftree}
\end{small}
\end{proof}

@@

					\AXC{$ $} \RightLabel{5}
					\UIC{$G(\gamma)$} \dottedLine
					\UIC{$\all \varphi. P(\varphi) \imp \varphi(\gamma)$} \RightLabel{$\all_E$}
					\UIC{$P(\lambda x. \neg \rho(x)) \imp \neg \rho(\gamma)$} \RightLabel{$\imp_E$}
				\BIC{$\neg \rho(\gamma)$} \RightLabel{$\neg_E$}      
	   \BIC{$\bot$} \RightLabel{$\neg_I^1$}
	   \UIC{$\neg\neg P(\rho)$} \RightLabel{$\neg\neg_E$}
	   \UIC{$[\subproof_4] \qquad P(\rho)$}
	   \end{prooftree}
\end{scriptsize}

@@

 \AXC{$\subproof_4$}\dashedLine
 \UIC{$P(\rho)$}
	      \AXC{Axiom 4} \dashedLine
        \UIC{$\all \varphi.[P(\varphi) \to \Box \; P(\varphi)]$} \RightLabel{$\all_E$}
	   	  \UIC{$P(\rho) \imp \nec P(\rho)$} \RightLabel{$\imp_E$}
		 \BIC{$\nec P(\rho)\scott$}
			\AXC{$ $} \RightLabel{3}
			\UIC{$\omega: P(\rho)$}
					\AXC{$ $} \RightLabel{2}
					\UIC{$\omega: G(\delta) $} \dottedLine
					\UIC{$\omega: \all \varphi. P(\varphi) \imp \varphi(\delta)$} \RightLabel{$\all_E$} 
					\UIC{$\omega: P(\rho) \imp \rho(\delta)$} \RightLabel{$\imp_E$}
				\BIC{$\omega: \rho(\delta) $} \RightLabel{$\imp_I^2$}
				\UIC{$\omega: G(\delta) \imp \rho(\delta) $} \RightLabel{$\all_I$}
				\UIC{$\omega: \all y. G(y) \imp \rho(y) $} \RightLabel{$\imp_I^3$}
				\UIC{$\omega: P(\rho) \imp (\all y. G(y) \imp \rho(y) )$} \RightLabel{$\nec_I$}
				\UIC{$\nec (P(\rho) \imp (\all y. G(y) \imp \rho(y) ))$} \RightLabel{$\imp_E$}
						\AXC{Axiom K} \RightLabel{$\imp_E$}
					\BIC{$\nec P(\rho) \imp \nec (\all y. G(y) \imp \rho(y) )$} \RightLabel{$\imp_E$}
			\BIC{$\nec (\all y. G(y) \imp \rho(y) )$} \RightLabel{$\imp_I^4$}
			\UIC{$\rho(\gamma) \imp \nec (\all y. G(y) \imp \rho(y) )$} \RightLabel{$\all_I$}
			\UIC{$\all \varphi. \varphi(\gamma) \imp \nec (\all y. G(y) \imp \varphi(y) )$} \RightLabel{$\wedge_I$}
		\BIC{$ G(\gamma) \wedge (\all \varphi. \varphi(\gamma) \imp \nec (\all y. G(y) \imp \varphi(y) ))$} \dottedLine
		\UIC{$\ess{G}{\gamma}$} \RightLabel{$ \imp_I^5$}
		\UIC{$G(\gamma) \imp \ess{G}{\gamma}$} \RightLabel{$\all_I$}
		\UIC{$\all x. G(x) \imp \ess{G}{x}$}
\end{prooftree}
\end{scriptsize}
